---
title: "P8105_hw3_cs3652"
author: "Chirag Shah"
date: '2018-10-15'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
```

##Problem 1

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  select(location_abbr = locationabbr, location_desc = locationdesc, year, topic, response, data_value) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response = factor(response, levels = ordered(c("Excellent", "Very good", "Good", "Fair", "Poor"))))
```

```{r}
brfss_data %>% 
  select(year, location_abbr, location_desc) %>% 
  filter(year == "2002") %>% 
  distinct(location_desc, location_abbr) %>% 
  group_by(location_abbr) %>% 
  summarize(locations_7 = n()) %>% 
  filter(locations_7 == "7")
```

```{r}
brfss_data %>% 
  select(year, location_abbr, location_desc) %>% 
  filter(year > 2001, year < 2011) %>% 
  group_by(location_abbr, year) %>% 
  distinct(location_desc, location_abbr) %>% 
  summarize(locations_per_state = n()) %>%
  ggplot(aes(x = year, y = locations_per_state, color = location_abbr)) +
  geom_line() +
  labs(
    title = "Number of Locations in Each State From 2002-2010",
    x = "year",
    y = "number of locations",
    caption = "data from the BRFSS dataset"
  ) +
  viridis::scale_color_viridis(
    name = "state", 
    discrete = TRUE
  ) + 
  ggthemes::theme_economist() +
  theme(legend.position = "bottom")
```

##col_types = "dccccccdddddccccccccc" %>% 

```{r}
brfss_data %>%
  spread(key = response, value = data_value) %>%
  select(year, Excellent, location_abbr) %>% 
  filter(!is.na(Excellent) & location_abbr == "NY" & (year == "2002" | year == "2006" | year == "2010")) %>%
  group_by(year) %>% 
  summarize(mean_proportion_excellent = mean(Excellent), sd_proportion_excellent = sd(Excellent)) %>% 
  knitr::kable()
```

```{r}
brfss_data %>% 
  select(year, location_abbr, response, data_value) %>% 
  group_by(year, location_abbr, response) %>% 
  summarize(avg_response = mean(data_value)) %>% 
  ggplot(aes(x = year, y = avg_response, color = location_abbr)) + 
  geom_point() +
  labs(
    title = "Average Proportion of Each Response Category in Each State Over Time",
    x = "year",
    y = "average proportion of response",
    caption = "data from BRFSS_2010"
  ) +
  viridis::scale_color_viridis(
    name = "state", 
    discrete = TRUE
  ) + 
  ggthemes::theme_economist() + 
  theme(legend.position = "bottom") +
  facet_grid(~ response) 
```

##Problem 2

```{r}

```

This problem uses the Instacart data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package (it’s called instacart).

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?
    Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
    Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
    Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).



