---
title: "P8105_hw3_cs3652"
author: "Chirag Shah"
date: '2018-10-15'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
fig.width = 6,
fig.asp = .6,
out.width = "90%"
)
```

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
library(patchwork)
library(hexbin)
```

##Problem 1

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  select(location_abbr = locationabbr, location_desc = locationdesc, year, topic, response, data_value) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response = factor(response, levels = ordered(c("Excellent", "Very good", "Good", "Fair", "Poor"))))
```

```{r}
brfss_data %>% 
  select(year, location_abbr, location_desc) %>% 
  filter(year == "2002") %>% 
  distinct(location_desc, location_abbr) %>% 
  group_by(location_abbr) %>% 
  summarize(locations_7 = n()) %>% 
  filter(locations_7 == "7")
```

```{r}
brfss_data %>% 
  select(year, location_abbr, location_desc) %>% 
  filter(year > 2001, year < 2011) %>% 
  group_by(location_abbr, year) %>% 
  distinct(location_desc, location_abbr) %>% 
  summarize(locations_per_state = n()) %>%
  ggplot(aes(x = year, y = locations_per_state, color = location_abbr)) +
  geom_line() +
  labs(
    title = "Number of Locations in Each State From 2002-2010",
    x = "year",
    y = "number of locations",
    caption = "data from the BRFSS dataset"
  ) +
  viridis::scale_color_viridis(
    name = "state", 
    discrete = TRUE
  ) + 
  ggthemes::theme_economist() +
  theme(legend.position = "bottom", legend.text = element_text(size = 6), legend.key.size = unit(2, "point"))
```

##col_types = "dccccccdddddccccccccc" %>% 

```{r}
brfss_data %>%
  spread(key = response, value = data_value) %>%
  select(year, Excellent, location_abbr) %>% 
  filter(!is.na(Excellent) & location_abbr == "NY" & (year == "2002" | year == "2006" | year == "2010")) %>%
  group_by(year) %>% 
  summarize(mean_proportion_excellent = mean(Excellent), sd_proportion_excellent = sd(Excellent)) %>% 
  knitr::kable()
```

```{r}
brfss_data %>% 
  select(year, location_abbr, response, data_value) %>% 
  group_by(year, location_abbr, response) %>% 
  summarize(avg_response = mean(data_value)) %>% 
  ggplot(aes(x = year, y = avg_response, color = location_abbr)) + 
  geom_point() +
  labs(
    title = "Average Proportion of Each Response Category in Each State Over Time",
    x = "year",
    y = "average proportion of response",
    caption = "data from BRFSS_2010"
  ) +
  viridis::scale_color_viridis(
    name = "state", 
    discrete = TRUE
  ) + 
  ggthemes::theme_economist() + 
  theme(legend.position = "bottom", legend.text = element_text(size = 6), legend.key.size = unit(2, "point")) +
  facet_grid(~ response) 
```

##Problem 2

```{r}
library(p8105.datasets)
data("instacart")

instacart_data = instacart %>% 
  janitor::clean_names()
```


To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):
`r nrow(instacart_data)` `r ncol(instacart_data)` observations: `r count(instacart_data)` distict users `r instacart_data %>% distinct(user_id) %>% count()` 
key variables: `user_id` `r instacart_data %>% select(user_id) %>% head(1)`
`product_name` `r instacart_data %>% select(product_name) %>% head(1)`
`department` `r instacart_data %>% select(department) %>% head(1)`

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart_data %>% 
  select(aisle) %>% 
  count()

instacart_data %>% 
  group_by(aisle) %>% 
  summarize(amount_ordered = n()) %>% 
  arrange(desc(amount_ordered)) %>% 
  select(aisle) %>% 
  head(1)
```

Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
```{r}
instacart_data %>% 
  group_by(aisle) %>% 
  summarize(amount_ordered = n()) %>% 
  ggplot(aes(x = aisle, y = amount_ordered)) + 
  geom_point() +
  labs(
     title = "Items ordered in each aisle",
     x = "aisle",
     y = "number of items ordered"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

    Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
    
```{r}
instacart_data %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(number_ordered = n()) %>% 
  arrange(desc(number_ordered)) %>% 
  group_by(aisle) %>% 
  top_n(n = 1) %>% 
  rename(most_popular_item = product_name) %>% 
  knitr::kable()
```

    Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart_data %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour) %>% 
  knitr::kable()
```

##Problem 3 

```{r}
library(p8105.datasets)
data(ny_noaa)
```

`r nrow(ny_noaa)`
`r ncol(ny_noaa)` 
`r count(ny_noaa)` observations
`r ny_noaa %>% distinct(id) %>% count()` distinct weather stations
key variables: id `r ny_noaa %>% select(id) %>% head(1)` 
date of the weather observation `r ny_noaa %>% select(date) %>% head(1)`
precipitation that occurred 
tmax and tmin variables give maximum and minimum temperatures
missing data values in this data set for prcp, tmin, and tmax `r ny_noaa %>% filter(is.na(tmax) | is.na(tmin) | is.na(prcp) | is.na(snow) | is.na(snwd)) %>% nrow()` 

```{r}
ny_noaa_data = ny_noaa %>%
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(prcp = prcp / 10, tmax = as.integer(tmax) / 10, tmin = as.integer(tmin) / 10)
  ##change to appropriate units since prcp was in tenths of mm and temperatures were in tenths of degrees celcius 
ny_noaa_data %>% 
  group_by(snow) %>% 
  summarize(freq_observed = n()) %>% 
  arrange(desc(freq_observed)) %>% 
  head(1)
```

Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
ny_noaa_data %>% 
  filter(tmax != "NA", (month == "01" | month == "07")) %>% 
  group_by(id, year, month) %>% 
  summarize(avg_tmax = mean(tmax)) %>%
  ggplot(aes(x = year, y = avg_tmax, color = id)) +
    geom_point() +
  scale_x_discrete(breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010)) +
  facet_grid(~month) +  
  labs(
      x = "year",
      y = "average temp (°C)",
      title = "Average Maximum Temperatures for January and July in New York Weather Stations Over Time",
      caption = "Data from NOAA"
    ) +
    viridis::scale_color_viridis(
      discrete = TRUE,
      name = "ID"
    ) +
    ggthemes::theme_economist() +
    theme(legend.position = "bottom", legend.text = element_text(size = 6), legend.key.size = unit(2, "point")) +
    guides(color = guide_legend(nrow = 20))
```

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
t_max_min_plot = ny_noaa_data %>% 
  filter(tmax != "NA" & tmin != "NA") %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
    geom_hex() + 
  labs(
      x = "max temperature (°C)",
      y = "min temperature (°C)",
      title = "Maximum and minimum temperatures for full dataset",
      caption = "Data from NOAA"
    ) + 
  ggthemes::theme_economist() +
  theme(legend.position = "bottom", legend.text = element_text(size = 6), legend.key.size = unit(2, "point"))

snowfall_plot = ny_noaa_data %>% 
  filter(snow != "NA" & snow > 0 & snow < 100) %>% 
  ggplot(aes(x = year, y = snow)) +
  geom_boxplot() +
    labs(
      x = "year",
      y = "snowfall amount (mm)",
      title = "distribution of snowfall by year",
      caption = "Data from NOAA"
    ) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 5))

t_max_min_plot / snowfall_plot
```


